---
title: "Patterns over Time"
---


# Overview of Trends

We start by plotting the overall number of attacks, along with the number of death and injury incurred. 


```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10}



total_number = GTD %>%
  group_by(iyear) %>%
  summarize(total_incidents = n())

totaldeath = GTD[ ,c("iyear", "nkill")] %>%
  group_by(iyear) %>%
  summarize(total_death = sum(nkill, na.rm = T))

totalwound = GTD[ ,c("iyear", "nwound")] %>%
  group_by(iyear) %>%
  summarize(total_wound = sum(nwound, na.rm = T))

trend2 = matrix(c(rep(0, 270)), nrow = 135, ncol =2)
trend2[ , 1] =  rep(total_number$iyear, 3)
trend2[ ,2][1:45] = total_number$total_incidents
trend2[ ,2][46:90] = totaldeath$total_death
trend2[ ,2][91:135] = totalwound$total_wound
colnames(trend2) = c("iyear", "total")
trend2 = cbind(trend2, matrix(c(rep(0, 135))))
trend2 = as.data.frame(trend2)
colnames(trend2) = c("iyear", "total", "category")

trend2[ ,3][1:45] = "Attacks"
trend2[ ,3][46:90] = "Deaths"
trend2[ ,3][91:135] = "Injuries"
colnames(trend2) = c("iyear", "total", "category")


trend2$category = as.factor(trend2$category)
trend2$iyear = as.numeric(trend2$iyear)
trend2$total = as.numeric(trend2$total)

plot1 = ggplot(trend2, aes(x = iyear, y = total, color = category)) +
  geom_line() +
  xlab("") +
  ylab("Total Number of Terror Attacks") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 18)) +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  ggtitle("Terrorist Incidents, 1970-2015") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_hc(bgcolor = "darkunica") +
  scale_fill_hc("darkunica")  +
  guides(fill=guide_legend(title=NULL)) +
  theme(legend.position="top")
plot1


```

We observed high number of incidents after 2010. Here we random picked four years to see whether there is any pattern within a year. According to the plot, there is no obvious peak for specific month for all years. This is reasonable because if the terrorism patterns are consistent over years, the police will stay alert during the days that attacks have a highly chance to happen.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10}
a <- GTD
a$iday[a$iday == 0] <- NA
a$date <- paste(a$iyear, a$imonth, a$iday, sep = "-")
a$date <- as.Date(a$date, "%Y-%m-%d")

b <- a %>% group_by(iyear, imonth) %>% summarise(count = n())
b$month <- as.factor(b$imonth)

b %>% 
  filter(iyear %in% c(2009, 2011, 2013, 2015)) %>% 
  ggplot(aes(x = month, y = count, group = 1)) + 
    geom_point(color = "orange") + 
    geom_line(color = "orange") + 
    facet_wrap(~iyear, ncol = 1, scale = "free") + 
    theme_hc(bgcolor = "darkunica") + 
    scale_fill_hc("darkunica") + 
    labs(x = "Month", y = "Number of Attacks") + 
    ggtitle("Global Trend of Terrorism Attacks, 1970-2015") + 
    theme(plot.title = element_text(size = 18, hjust = 0.45), 
          strip.background = element_blank(),
          strip.text = element_text(color = "white"))
```










Here we break down the total number of attacks per year by the type of the attack. We note that bombings have increased quite a lot relative to other types of attacks.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10}
newdata = GTD[ ,c("iyear","attacktype1_txt", "suicide", "weaptype1_txt", "targtype1_txt")]
attack = newdata %>%
  group_by(iyear, attacktype1_txt) %>%
  summarize(total = n())
attack$attacktype1_txt = as.factor(attack$attacktype1_txt)
levels(attack$attacktype1_txt)[levels(attack$attacktype1_txt) == "Hostage Taking (Barricade Incident)"] <- "Barricade Incident"
levels(attack$attacktype1_txt)[levels(attack$attacktype1_txt) == "Hostage Taking (Kidnapping)"] <- "Kidnapping"
levels(attack$attacktype1_txt)[levels(attack$attacktype1_txt) == "Hostage Taking (Kidnapping)"] <- "Kidnapping"
levels(attack$attacktype1_txt)[levels(attack$attacktype1_txt) == "Facility/Infrastructure Attack"] <- "Facility/Infrastructure"

colnames(attack)[2] = 'Attack'

attack_1 <- attack %>%
  filter(Attack != "Unknown") %>%
  ggplot(aes(x = iyear, y = total, fill = Attack)) +
  geom_bar(stat = "identity") +
  theme_hc(bgcolor = "darkunica") +
  scale_fill_hc("darkunica") +
  xlab("") +
  ylab("") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 18)) +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  ggtitle("Types of Terror Attacks") +
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(fill=guide_legend(title=NULL)) +
  theme(legend.position="top")
attack_1

```

We do the same for the types of targets that were victim to these attacks. As the plot shows, private citizens & property has experienced the largest relative increase.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10}
target = newdata %>%
  group_by(iyear, targtype1_txt) %>%
  summarize(total = n())

target$targtype1_txt[target$targtype1_txt == 'Government (Diplomatic)'| target$targtype1_txt == 'Government (General)'] = "Government"
target$targtype1_txt[target$targtype1_txt == 'Airports & Aircraft'] = "Transportation"
target$targtype1_txt[target$targtype1_txt == 'Business'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Educational Institution'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Food or Water Supply'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Journalists & Media'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'NGO'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Journalists & Media'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Religious Figures/Institutions'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Telecommunication'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Terrorists/Non-State Militia'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Unknown'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Violent Political Party'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Maritime'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Abortion Related'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Tourists'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Utilities'] = "Other"
target$targtype1_txt[target$targtype1_txt == 'Private Citizens & Property'] = "Citizens&Property"

target = target %>%
  group_by(iyear, targtype1_txt) %>%
  summarise(total = sum(total))
target$targtype1_txt = as.factor(target$targtype1_txt)

target_1 <- target %>%
  ggplot(aes(x = iyear, y = total, fill = reorder(targtype1_txt, desc(total)))) +
  geom_bar(stat = "identity") +
  theme_hc(bgcolor = "darkunica") +
  scale_fill_hc("darkunica") +
  xlab("") +
  ylab("") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 18)) +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  ggtitle("Types of Targets") +
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(fill=guide_legend(title=NULL)) +
  theme(legend.position="top")
target_1
```

Interstingly, suicide attacks experienced "peak popularity" during the height of the Iraq War in 2004 - 2006.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10}

suicide = newdata %>%
  group_by(iyear, suicide) %>%
  summarize(total = n())

a = suicide %>%
  filter(suicide == 1)
b = suicide %>%
  filter(suicide == 0)
b = b[ , c(1,3)]
a = a[ , c(1,3)]
suicide2 = left_join(b, a, by = c("iyear"="iyear"))
suicide2[is.na(suicide2)] <- 0
suicide2 %>%
  group_by(iyear) %>%
  summarize(rate = total.y / total.x * 100) %>%
  ggplot(aes(x = iyear, y = rate)) +
  geom_line(color = "orange") + 
  xlab("Year") +
  ylab("Percent (%)") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 18)) +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  ggtitle("% of Suicide Attacks, 1970-2015") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_hc(bgcolor = "darkunica") +
  scale_fill_hc("darkunica") 



```

