---
title: "Terrorist Organizations"

---
# Patterns Across Terrorist Organizations



Finding that top 10 terrorist organization already account for 20% of the total attacks.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10}
# gang data
library(dplyr)
gangs <- GTD %>%
  group_by(gname) %>%
  summarise(num_count = n(),
            num_killed = sum(nkill, na.rm = TRUE),
            num_wounded = sum(nwound, na.rm = TRUE)) %>%
  arrange(desc(num_count)) %>%
  mutate(rank_count = 1:length(unique(gname)), 
         fraction_count = cumsum(num_count) / sum(num_count, na.rm = TRUE),
         fraction_killed = cumsum(num_killed) / sum(num_killed, na.rm = TRUE),
         fraction_wounded = cumsum(num_wounded) / sum(num_wounded, na.rm = TRUE))

library(tidyverse)
gang_plotdata <- gather(gangs, key = harm, value = number, c(6:8))

gang_discriptive <- gangs %>% 
  filter(rank_count %in% c(1, 11, 51, 101)) %>%
  select(gname, rank_count, fraction_count, fraction_killed, fraction_wounded)

# plot
library(ggthemes)
library(plotly)
ggplot(gang_plotdata, aes(x = rank_count, y = number, color = harm)) +
  geom_line(size = 1) +
  geom_hline(yintercept = 0.45, color = "white", lty = 2) +
  geom_hline(yintercept = 0.65, color = "white", lty = 2) +
  geom_hline(yintercept = 0.87, color = "white", lty = 2) +
  annotate("text", x = 3500, y = 0.5, size = 6, adj = 1, 
           family = "serif", color = "white", label = c("Baseline")) +
  annotate("text", x = 3500, y = 0.71, size = 6, adj = 1, 
           family = "serif", color = "white", label = c("10 top gangs")) +
  annotate("text", x = 3500, y = 0.93, size = 6, adj = 1, 
           family = "serif", color = "white", label = c("100 top gangs")) +
  theme_hc(bgcolor = "darkunica") +
  ggtitle("Fraction of Harm Done") +
  labs(x = "Number of Gangs", y = "Count of Bad Consequences") +
  scale_fill_hc("darkunica") +
  theme(plot.title = element_text(size = 18, hjust = 0.45), 
          strip.background = element_blank(),
          strip.text = element_text(color = "white"))

```


So we want to focus more on the top terrorist organizations to draw more conclusions. In order to determine who are the top gangs, we made an assumption to define the importance of a particular organization: 

1) It is very active (has large number of attacks) 
2) Its attacks have high fatality rate

Here is the listing of these top offenders:

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=15}
gang_rank <- GTD %>%
  group_by(gname) %>%
  summarise(total_num_attacks = n(),
            death_ratio = round(mean(nkill, na.rm = TRUE), digits = 1),
            wounded_ratio = round(mean(nwound, na.rm = TRUE)), digit = 1) %>%
  mutate(active_rank = rank(-total_num_attacks, na.last = TRUE),
         deathratio_rank = rank(-death_ratio, na.last = TRUE),
         woundedratio_rank = rank(-wounded_ratio, na.last = TRUE))

library(DT)
winners <- rbind(filter(gang_rank, active_rank == 2),
                 filter(gang_rank, deathratio_rank == 1),
                 filter(gang_rank, woundedratio_rank == 1))
winners$winners <- c("Most frequent activities", "Top fatality rate", "Top injury rate")
winners <- winners %>%
  select(c(9, 1:4)) %>%
  rename(Winners = winners, Gang_Name = gname, Frequency = total_num_attacks, 
         Fatality_Rate = death_ratio, Injury_Rate = wounded_ratio)
datatable(winners, style = "uikit", width = 20)
```

We can see that these three winner gangs perform their attack in different styles. Taliban has the most attacks throughout the years; Dissident Military Members of Northern Tribal Group has just one attack but killed 150 people with none survived; Aum Shinri Kyo has several attacks, wounding many people but not causing heavy fatal consequences.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=15}
top_gang <- gang_rank %>%
  filter(active_rank <= 50, deathratio_rank <= 500, woundedratio_rank <= 508) %>%
  arrange(active_rank) %>%
  select(1:4)
datatable(top_gang, width = 20)
```

We filter out 9 gangs which qualifies for "big gangs". We find that although they only account for 1% of total terrorist attacks, they are responsible for 3% of total death and 2% of total wounded.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10}
total <- c(nrow(GTD), 
           sum(GTD$nkill, na.rm = TRUE), 
           sum(GTD$nwound, na.rm = TRUE))

top_gname <- top_gang$gname
top_gang_record <- filter(GTD, gname == top_gname)
top <- c(nrow(top_gang_record), 
         sum(top_gang_record$nkill, na.rm = TRUE), 
         sum(top_gang_record$nwound, na.rm = TRUE))

big_small <- data.frame(feature = c("total_attack", "total_death", "total_wounded"),
                                    total = total, top = top, percentage = round(top / total, digits = 2))
big_small$percentage <- paste(big_small$percentage * 100, "%")
big_small <- gather(big_small, key = scope, value = count, c(2:3))
big_small$percentage[4:6] <- NA

ggplot(big_small, aes(x = feature, y = count, fill = scope)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = percentage)) +
  theme_hc(bgcolor = "darkunica") +
  scale_fill_hc("darkunica") +
  ggtitle("Percentage of Harm that Top Gangs are Responsible for") +
  labs(x = "Type of Consequence", y = "Count") +
  theme(plot.title = element_text(size = 18, hjust = 0.45), 
          strip.background = element_blank(),
          strip.text = element_text(color = "white"))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10}
library(devtools)
library(leaflet)


keep <- c("eventid", "iyear", "imonth", "iday", "country", "country_txt", "latitude", "longitude", "attacktype1_txt", "targtype1_txt", "gname", "weaptype1_txt", "nkill", "nwound")
top_gang_record <- top_gang_record[ , keep]
top_gang_record$gname <- as.factor(top_gang_record$gname)
top_gang_record$severity <- sum(top_gang_record$nkill + top_gang_record$nwound)
```

We can see that these top gangs each has a territory of its own. 
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10}
library(RColorBrewer)
pal = colorFactor("Set1", domain = top_gang_record$gname)
color_offsel1 <- pal(top_gang_record$gname)

content <- paste("Gang Name:", top_gang_record$gname,"<br/>",
                 "When:", top_gang_record$iyear,"<br/>",
                 "Where:", top_gang_record$country_txt,"<br/>",
                 "Fatality:", top_gang_record$nkill,"<br/>",
                 "Injury:", top_gang_record$nwound, "<br/>")       

gangs_map <- leaflet(top_gang_record) %>% 
     addTiles('http://{s}.basemaps.cartocdn.com/dark_all/{z}/{x}/{y}.png') %>% 
  addCircles(col="orange") %>%
  addCircles(color = color_offsel1) %>%
  addCircles(color = color_offsel1, popup = content) %>%
  addCircleMarkers(color = color_offsel1, 
                       popup = content, radius = ~ log(nkill, 1.2), fillOpacity = 0.7)

gangs_map
```

With this plot we see a clear expansion of ISIS, from limited to Iraq to spreading to Afraica and eventually France.
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10}
ISIS <- GTD %>% filter(gname == c("Islamic State of Iraq and the Levant (ISIL)", 
                                 "Islamic State of Iraq (ISI)"))

content_2 <- paste("Gang Name:", ISIS$gname,"<br/>",
                 "When:", ISIS$iyear,"<br/>",
                 "Where:", ISIS$country_txt,"<br/>",
                 "Fatality:", ISIS$nkill,"<br/>",
                 "Injury:", ISIS$nwound, "<br/>")      


ISIS_map <- leaflet(ISIS) %>% 
     addTiles(group = "base",  'http://{s}.basemaps.cartocdn.com/dark_all/{z}/{x}/{y}.png') %>%
  
  addCircles(group = "Before 2012", data = filter(ISIS, iyear <= 2012), 
             lng = ~longitude, lat = ~latitude, col="white") %>%
  # addCircleMarkers(color = "white", popup = content_2, 
                   # radius = ~ log(nkill, 1.2), fillOpacity = 0.7) %>%
  
  addCircles(group = "2013", data = filter(ISIS, iyear == 2013), 
             lng = ~longitude, lat = ~latitude, col="yellow") %>%
  # addCircleMarkers(color = "yellow", popup = content_2, 
                   # radius = ~ log(nkill, 1.2), fillOpacity = 0.7) %>%
  
  addCircles(group = "2014", data = filter(ISIS, iyear == 2014), 
             lng = ~longitude, lat = ~latitude, col = "orange") %>% 
  # addCircleMarkers(color = "orange", popup = content_2, 
                   # radius = ~ log(nkill, 1.2), fillOpacity = 0.7) %>%
  
  addCircles(group = "2015", data = filter(ISIS, iyear == 2015), 
             lng = ~longitude, lat = ~latitude, col = "red") %>%
  # addCircleMarkers(color = "red", popup = content_2, 
                   # radius = ~ log(nkill, 1.2), fillOpacity = 0.7) %>%
  
   addLayersControl(
    baseGroups = c("base"),
    overlayGroups = c("Before 2012", "2013", "2014", "2015"),
    options = layersControlOptions(collapsed = TRUE))

ISIS_map
```

